---
title: "Comparison of Gene Weights"
date: "`r strftime(Sys.time(), '%Y-%m-%d')`"
params:
   rmd: "combined_weights_summary.Rmd"
output:
  html_document:
    df_print: kable
    toc: true
    toc_depth: 2
---

# Overview

In this file, different versions of the gene weightings (generated via different
gene-phenotype associations methods or parameters and different methods for generating
combined weights from the inidivual sources) are compared.

Naming Conventions:

- **dataset**: `XX_pearson_max` -- maximum feature-pheno pearson correlation used as weight
- **field**: `max_score`: when combining weights from different sources, the maximum
  value is used.

(Note: the handling and presentation of dataset sources, methods used to measure
gene-phenotype associations, and methods for collapsing individual dataset weights needs
to be improved..)

# Setup

```{r, message = FALSE}
library(knitr)
library(gridExtra)
library(tidyverse)

options(stringsAsFactors = FALSE)
options(digits = 3)

opts_chunk$set(fig.width = 20, fig.height = 11.24, dpi = 192, retina = 1)
```

# Feature Weights

```{r include = FALSE}
# debug
if (snakemake@config$dev_mode$enabled) {
  if (!dir.exists('/rda/nih/fw/reports')) {
    dir.create('/rda/nih/fw/reports', mode='0755', recursive=TRUE)
  }
  save.image('/rda/nih/fw/reports/feature_weights_summary.rda')
}
```

```{r}
if (snakemake@wildcards$feature_level == 'gene_sets') {
  feat_ids <- c('collection', 'gene_set')
  feat_id_indices <- 1:2
} else {
  feat_ids <- 'symbol'
  feat_id_indices <- 1
}
```

```{r load_feature_weights}
# load feature weights
gene_weight_files <- c(snakemake@input$indiv_weights, snakemake@input$combined_weights)

# drop redundant files due to expand
gene_weight_files <- unique(gene_weight_files)

wts_list <- lapply(gene_weight_files, read_tsv, col_types = cols())
names(wts_list) <- gene_weight_files

sapply(wts_list, nrow)
sapply(wts_list, ncol)
```

## Distribution of Feature Weghts

```{r, results = 'asis', dpi = 128}
for (dataset_id in names(wts_list)) {
  cat(sprintf("\n### %s\n\n", dataset_id))

  dataset <- wts_list[[dataset_id]]

  plts <- list()

  for (feat_wt in colnames(dataset)[-feat_id_indices]) {
    plt_title <- sprintf("%s - %s", dataset_id, feat_wt)

    if (grepl('auc', feat_wt)) {
      plt_color <- '#e31a51'
    } else if(grepl('ic50', feat_wt)) {
      plt_color <- '#4fe319'
    } else {
      plt_color <- '#194fe3'
    }

    plt_dat <- data.frame(x = dataset[[feat_wt]])

    plts[[plt_title]] <- ggplot(plt_dat, aes(x)) +
      geom_density(fill = plt_color) +
      xlab('weight') +
      ggtitle(plt_title)
  }

  grid.arrange(grobs = plts, ncol = min(length(plts), 3))

  cat('\n')
}
```

# Top features

```{r, results = 'asis'}
for (dataset_id in names(wts_list)) {
  cat(sprintf("## %s\n", dataset_id))

  wts <- wts_list[[dataset_id]] %>%
    pivot_longer(-feat_ids, names_to = 'method', values_to = 'score')

  for (score_method in unique(wts$method)) {
    cat(sprintf("### %s\n", score_method))

    wts %>%
      filter(method == score_method) %>%
      select(-method) %>%
      arrange(desc(score)) %>%
      head(15) %>%
      kable() %>%
      print()

    cat('\n')
  }
}
```

```{r include = FALSE}
if (snakemake@config$dev_mode$enabled) {
  save.image('/rda/nih/fw/combined_weights_summary.rda')
}
```


# Session Information

```{r}
sessionInfo()
```

